Saint Helena Island Implementation

### 2025-01-13 Implementation doc/Draft1.Rmd
Add files inside Implementation doc

This document describes the configuration of StrathE2E for the Saint Helena Island off the South Mid-Atlantic Ridge and its parameterisation to enable stationary state fitting for both the baseline period (2010-2019) and future projections (2020-2070). These represent contrasting periods of environmental conditions.

@@

@@

@@

The migratory fishes in the region are part of a broader Atlantic stock and perform transatlantic movements. These species utilize multiple spawning grounds and are not present in the Atlantic year-round. Key migratory species making seasonal transits through Saint Helena waters include skipjack tuna (*Katsuwonus pelamis*), bigeye tuna (*Thunnus obesus*), and yellowfin tuna (*Thunnus albacares*), along with other tuna-like species and sharks. The data on the biomass of these migratory fish stocks, the proportion entering the Saint Helena Island EEZ, and the timing of their migrations are derived from literature and stock assessments, as detailed below.

@@

Saint Helena likely serves as a feeding ground for sub-adult fish, which migrate to spawning grounds upon reaching maturity (Wright et al., 2020). Bigeye tuna and yellowfin tuna primarily spawn in tropical waters, such as the Gulf of Guinea and the Gulf of Mexico, during the summer (Carleton et al., 2010). In the Saint Helena EEZ, yellowfin tuna landings exhibit no significant seasonality, while bigeye tuna are more prevalent between September and December, skipjack from December to June, and albacore from July to September (Wright et al., 2020).

Recent tagging studies on yellowfin tuna show that they remain in Saint Helena waters for extended periods, ranging from six months to a year, during which they feed and grow (Wright et al., 2020). They are typically recruited into the fishery between February and April each year and leave the population upon reaching a length of 120 cm (Wright et al., 2020). Yellowfin tuna reach an average size of 96 to 120 cm at Cardno Seamount in May/June, suggesting migration out of the EEZ at this size (Wright et al., 2020; Wright et al., 2021). Additionally, yellowfin tuna exhibit high fidelity to the area and remain in St. Helena waters for extended periods due to elevated productivity levels associated with the island and seamounts (White et al., 2007; Wright et al., 2021; Wright et al., 2019). Recapture rates of conventionally tagged fish ranged from 12% (2017) to 30% (2015), with an overall average recapture rate of 19%.

The immigration period is set from September 1st to November 30th, and the emigration period from June 1st to July 31st, based on the months of highest presence of migratory fish described above. We also assumed that 20% of the population remains within the Saint Helena EEZ even during the immigration period.

To estimate the proportion of migratory stocks entering the Saint Helena domain, we conducted a rough estimation of their biomass in both the Atlantic Ocean and the Saint Helena EEZ.

Among the tuna species in the region, yellowfin tuna is the only one with a biomass assessment as, to date, tagging was insufficient to assess the local biomass of skipjack, albacore, and bigeye tuna within St. Helena's EEZ (Wright et al., 2020). Biomass estimates from tagging recoveries by Wright et al. (2020) between 2016 and 2019 ranged from 1,370 to 2,935 tons, with an average of 1,916 tons of yellowfin tuna biomass across three key fishing grounds in Saint Helena. The ICCAT target exploitation rate (FMSY) for yellowfin tuna is 16% (ICCAT, 2019), and the local exploitation rate aligns with this, given the current TAC of 300 tons. This suggests a yellowfin biomass of approximately 1,875 tons, which closely matches the 1,916-ton estimate. Therefore, we used the baseline yellowfin biomass of 1,916 tons.

For skipjack tuna, St. Helena typically accounts for 0.42% of the East Atlantic skipjack stock landed within its EEZ (Wright et al., 2020). Our calculations are based on the proportion of catch in the Eastern Atlantic during the baseline period, utilizing biomass estimates from Sea Around Us (SAU) stock assessments, which primarily rely on ICCAT data and the CMSY method (Pauly et al., 2020). Covering the period from 2010 to 2018, we averaged this proportion and applied to the TAC of 750 tons per year in St. Helena to obtain a rough estimate of the biomass.

For bigeye tuna, the most recent ICCAT assessment (ICCAT, 2021) indicates that the stock is overfished but not currently undergoing overfishing, following international landing reductions in 2018. The TAC for the UK overseas territories was initially set at 1,000 tons, though local authorities in Saint Helena questioned whether local bigeye stocks could sustain such a catch (Wright et al., 2020). Consequently, in 2020, the TAC was reduced to 600 tons. To estimate bigeye tuna biomass, we applied the ICCAT target exploitation rate of 19% to the most recent TAC.

The maximum sustainable yield (MSY) for albacore was 25,900 tons, with a relative fishing rate of F2014/FMSY at 0.54. ICCAT set the South Atlantic TAC for albacore at 24,000 tons (ICCAT, 2021), with the local TAC for St. Helena at 100 tons (REC 16-07), which accounts for 0.42% of the total. During the same period, the biomass was estimated at approximately 132,500 tons (ICCAT, 2020). Applying the 0.42% catch proportion to the biomass provides an estimate of the annual local albacore biomass.

Beyond tuna species, other significant migratory species in the region include swordfish (*Xiphias glaudis*). Although swordfish landings and sales are prohibited in St. Helena due to high mercury levels (TAC St. Helena, 2020), ICCAT sets a TAC of 25 tons for the UK overseas territories. The South Atlantic TAC for swordfish is 14,000 tons, with St. Helena's share corresponding to 0.18%. The estimated local swordfish biomass was 57,474 tons. Similar estimates for wahoo (*Acanthocybium solandri*), blue marlin (*Makaira nigricans*), and shortfin mako shark (*Isurus oxyrinchus*) were calculated using SAU data and catch-to-biomass ratios for the Atlantic stock between 2010 and 2018, and these were applied to the current TAC or reported catches in St. Helena where no TAC exists. For the galapogos shark *Carcharhinus galapagensis* and the white marlin *Kajikia albida*, we used the average local biomass estimates from the SAU assessment covering the period from 2010 to 2018.

Beyond tuna species, other significant migratory species in the region include swordfish (*Xiphias glaudis*). Although swordfish landings and sales are prohibited in St. Helena due to high mercury levels (TAC St. Helena, 2020), ICCAT sets a TAC of 25 tons for the UK overseas territories. The average swordfish catch in the South Atlantic was 14,000 tons, and applying the same catch-to-biomass ratios we obtained a local swordfish biomass of 159.29 tons. Similar estimates for wahoo (*Acanthocybium solandri*), blue marlin (*Makaira nigricans*), and shortfin mako shark (*Isurus oxyrinchus*) were calculated using SAU data and catch-to-biomass ratios for the Atlantic stock between 2010 and 2018, and these were applied to the current TAC or reported catches in St. Helena where no TAC exists. For the galapagos shark *Carcharhinus galapagensis* and the white marlin *Kajikia albida*, we used the average local biomass estimates from the SAU assessment covering the period from 2010 to 2018.

For whale sharks (*Rhincodon typus*), which are not fished but are seasonally abundant in Saint Helena from December to May each year (Perry et al., 2020), 277 individual sharks were identified through photographic identification between 2013 and 2019 (Perry et al., 2020). We used this number as a reference, estimating their biomass by multiplying the average weight from FishBase.

Our estimates resulted in an annual migratory fish biomass of 22,913 tons within the Saint Helena EEZ. Our biomass estimates for the 12 main stocks in the Atlantic region occurring around Saint Helena totaled 3,275,099.7 tons. The proportion of the migratory stock entering the model domain was calculated to be 0.7%. For tuna species alone, we estimated a proportion of 0.41%, consistent with the reported catch proportions from Saint Helena relative to total Atlantic catches. Figure \@ref(fig:migratory)) shows the dominant species in the Saint Helena Island EEZ in terms of biomasses compiled here.

```{r migratory, echo=FALSE, fig.align='center', fig.cap="Proportion of biomasses by species of the assessed migratory guild in the Saint Helena EEZ", out.width = '50%'}
knitr::include_graphics("img/migratory.png")
```

```{r migrant_cap}

tab_migrant_cap <- table_nums(name = "tab_migrant", 
    caption = "Biological event timing parameters, constant across the time periods. The data are processed in the model setup to calculate the immigration flux parameters needed in the ecology model. Spawning and recruitment durations were established assuming a month with 30 days.")
```

`r table_nums("tab_migrant")`

```{r tab_migrant, fig.cap = tab_migrant_cap}

data_frame(Parameter = model[["data"]][["biological.events"]][["Description"]] %>% 
                        str_replace_all("_", " "),
           Value = model[["data"]][["biological.events"]][["Value"]]) %>% 
  .[1:17,] %>% 
  gt() %>%
  fmt_number(
    decimals = 0,
    use_seps = FALSE
  )

\newpage

`r table_nums("tab_migrant", display = "cite")` Continued.

```{r tab_migrantb}

data_frame(Parameter = model[["data"]][["biological.events"]][["Description"]] %>% 
                        str_replace_all("_", " "),
           Value = model[["data"]][["biological.events"]][["Value"]]) %>% 
  .[18:26,] %>% 
  gt() %>%
  fmt_number(
    decimals = 2,
    use_seps = FALSE
  )
```

@@

-   Temperature, vertical mixing coefficients, volume fluxes, and boundary nutrient, detritus and phytoplankton concentrations from outputs of a NEMO-ERSEM coupled hydro-geochemical model run at SSP370 with a 2010-2019 historical/future split [@yool13].
-   Surface shortwave radiation, surface air temperature, and freshwater volume outflows from HadGEM2-ES model output [@jones11] used to force the NEMO-MEDUSA coupled hydro-geochemical model mentioned above [@yool13].
-   River nitrate and ammonia concentrations estimated from GRQA database (Global River Water Quality Archive, @virro21)
-   Atmospheric deposition of nitrate and ammonia from ISIMIP3a ([Inter-Sectoral Impact Model Intercomparison Project](https://data.isimip.org/10.48364/ISIMIP.759077.2); [@yang22])
-   Remote sensing data products on Suspended Particulate Matter (Globcolour L3b; <ftp://ftp.hermes.acri.fr/GLOB/merged/month/>).
-   Habitat disturbance due to tidal currents and waves from..
-   Wave height, period, and direction from the ERA-5 reanalysis monthly means accessed through [CDS](https://cds.climate.copernicus.eu/cdsapp#!/dataset/10.24381/cds.f17050d7?tab=overview) for 1980-2021 [@Hersbach19].

Details of how these data were processed are given below, supported by the [nemomedusR](https://jack-h-laverick.github.io/nemomedusR/) packages.

### Vertical mixing coefficients between the upper and lower layers of the deep zone: {.unnumbered}

Vertical diffusivity from the NEMO-MEDUSA coupled hydro-geochemical model output [@yool13] was interpolated for each grid cell at the SDepth m boundary depth between the shallow and deep layers of the offshore zone. These values were summarised as monthly averages into period-specific climatological annual cycles of data for the simulation periods.

@@

Derived by monthly averaging values at grid points within the inshore and vertical layers of the offshore zones from the NEMO-MEDUSA coupled hydro-geochemical model output [@yool13], weighted by grid point volumes. Values were summarised into period-specific climatological annual cycles of data for the simulation periods.

@@

Derived from HadGEM2-ES model output [@jones11] which forces the NEMO-MEDUSA model run used throughout our implementation. Monthly mean values were summarised into a climatological annual cycle of data for the simulation periods.

@@

Sourced from the "histsoc" files for a 1901 - 2021 hindcast as monthly averages [@yang22], available from [CDS](https://cds.climate.copernicus.eu/cdsapp#!/dataset/10.24381/cds.f17050d7?tab=overview). Monthly values were summarised into climatological annual cycles of monthly oxidised and reduced nitrogen deposition rates extracted for 2010-2019. Projections under different SSPs are also available to 2100.

@@

Freshwater inflow derived from HadGEM2-ES model output [@jones11] which forces the NEMO-MEDUSA model run used throughout our implementation. Monthly values were summaries into a climatological annual cycle of data for all periods.

@@

Monthly averaged daily inflow and outflow volume fluxes derived by integrating daily mean velocities directed perpendicular to transects along the model domain boundary at grid points in each depth layer along transects through outputs from the NEMO-MEDUSA coupled hydro-geochemical model output [@yool13]. Monthly averaged daily inflow volume fluxes then divided by the volume of the receiving layer in the model domain to estimate a daily flushing rate. Period-specific climatological annual cycles of data used for simulation periods.

@@

NEMO-MEDUSA outputs included phytoplankton and suspended detritus, as well as Dissolved Inorganic Nitrogen (DIN). We calculated the depth-averaged concentrations for pixels within the shallow and deep layers of StrathE2E. We then sampled the pixels using the same transects around the model domain as for sampling volume fluxes. Only transects where water flowed into the model domain were sampled, and the average concentration of inflowing waters for target variables was calculated weighting by the flow rate across a transect and the cross-sectional area represented by a transect (average depth and length). Concentrations were then averaged into climatological annual cycles for all periods.

The GRQA (@virro21) database was utilized, providing nitrate and ammonium concentrations in mg/l from continental and global observation datasets. Data were filtered for Brazilian rivers along the SBS coastline, and those inflowing to the inshore zone were selected. Outliers were identified using the IQR method. The resulting values were then transformed into monthly concentrations in mMN.m^-3^.

@@

@@

@@

```{r fleets, echo=FALSE, fig.cap="Fishing fleets defined for the StrathE2E Saint Helena EEZ model, their operational descriptions, target-species and discards levels.", out.width = '110%'}
knitr::include_graphics("img/Fleets.png")
```

@@

effort_tab <- effort_table %>%
  gt() %>%
  tab_options(
    table.font.size = 11,
    column_labels.font.weight = "bolder"
  ) %>%
  cols_align(
    align = "center"
  ) %>%
  cols_label(
    Fleets = "Fleets StrathE2E Saint Helena",
    `Effort (s.year-1)` = html("Effort (s.year<sup>-1</sup>)"),
    `Effort (s.m2-1.day-1)` = html("Effort (s.m<sup>2</sup>.day<sup>-1</sup>)"),
  )

@@

According to Wright et al. (2020), the fisheries around St. Helena can be divided into two zones: the inshore area, extending within 30 nm of the island, and the offshore area, from 30 to 200 nm. The offshore zone encompasses key fishing grounds around the Bonaparte (west) and Cardno (north) seamounts, where migratory species are typically fished in the upper water column at depths of 75 to 200 meters (Carleton et al., 2010). While yellowfin and skipjack tuna are commonly caught in both inshore and offshore areas, bigeye and albacore tuna are mainly found in the deeper waters near the island and the seamounts (Wright et al., 2020). Thus, the pole-and-line fleet is divided into 31.66% offshore (seamounts) and 68.34% inshore. Most of the inshore tuna fishing occurs over coarse sediment (gravel), bedrock, and some sand.

In contrast, handline fishing for grouper and pot fishing for lobsters are confined to the inshore area and are not present around the Bonaparte and Cardno seamounts (Collins et al., 2012; Riley et al., 2020). These activities are associated with inshore rock substrates (Rees et al., 2016). Baitfish (Decapterus spp.) are caught with nets at the surface in waters 10 to 20 meters deep (Carleton et al., 2010), usually near sandy and rocky inshore beaches (Edwards, 1990).

Using this information and landings data categorized by area (from Wright et al., 2020), we allocated the fishing effort per habitat as outlined in Table X.

@@

Calculating the production involved compiling and adjusting biomass values for the model area and converting them to molar nitrogen mass following @greenstreet96. Subsequently, we convert the biomass for each species into gross production using PB ratios and converted them to molar nitrogen mass following @greenstreet96. The PB ratios sourced was also used as target data. For birds, cetaceans, and pinnipeds, net production was calculated using the relationship net production = 0.6 \* gross production.

@@

tab <- production_tab %>%
  gt() %>%
  fmt_number(
    columns = c(`P/B`, `mMN convertion`, `Production(t)`),
    decimals = 2,
    use_seps = FALSE
  ) %>%
  fmt_number(
    columns = c(`Production (mMN.m-2)`),
    decimals = 6,
    use_seps = FALSE
  ) %>%
  tab_options(
    table.font.size = 11,
    column_labels.font.weight = "bolder",
  ) %>%
  cols_align(
    align = "center"
  ) %>%
  cols_label(
    Guild = "Guild",
    `P/B` = "P/B",
    `Production(t)` = "Production(t)",
    `mMN convertion` = "mMN convertion",
    `Production (mMN.m-2)` = html("Production (mMN.m<sup>-2</sup>)")
  )
tab
```

@@

By multiplying the estimated number of individuals by the average weight of each species, as provided by BirdLife International (2024) and Sealife (2024), we obtained an approximate seabird biomass of 9.82 tons, equating to X t/km². To obtain the production per year we used the P/B ratio for seabirds from the Ecopath model of Capitani et al. (2021) based on Rocas Atoll, another tropical reef ecosystem in the Southwest Atlantic.

@@

@@

@@

## Diet proportions {-}

@@

### Seabirds {-}

@@

tab <- diet_tab %>%
  gt() %>%
  tab_options(
    table.font.size = 11,
    column_labels.font.weight = "bolder",
  ) %>%
  cols_align(
    align = "center"
  ) 
tab

@@

`r table_nums("list_species")`

>>>>

This document describes the configuration of StrathE2E for the Saint Helena Island off the South Mid-Atlantic Ridge and its parameterisation to enable stationary state fitting for both the baseline period (2010-2019) and future projections (2020-2069).
These represent contrasting periods of environmental conditions.

@@

```{r island, echo=FALSE, out.width = '100%'}
knitr::include_graphics("img/island.png")
```

@@

```{r event_cap}

tab_event_cap <- table_nums(name = "tab_event", 
    caption = "Biological event timing parameters, constant across the time periods. The data are processed in the model setup to calculate the immigration flux parameters needed in the ecology model. Spawning and recruitment durations were established assuming a month with 30 days.")
```

`r table_nums("tab_event")`

```{r tab_migrant, fig.cap = tab_event_cap}

data_frame(Parameter = model[["data"]][["biological.events"]][["Description"]] %>% 
                        str_replace_all("_", " "),
           Value = model[["data"]][["biological.events"]][["Value"]]) %>% 
  .[1:16,] %>% 
  gt() %>%
  fmt_number(
    decimals = 0,
    use_seps = FALSE
  )

@@

The migratory fishes in the region are part of a broader Atlantic stock and perform transatlantic movements, utilizing multiple spawning grounds.
Key migratory species making seasonal transits through Saint Helena waters include skipjack tuna (*Katsuwonus pelamis*), bigeye tuna (*Thunnus obesus*), and yellowfin tuna (*Thunnus albacares*), along with other tuna-like species and sharks.
The data on the biomass of these migratory fish stocks, the proportion entering the Saint Helena EEZ, and the timing of their migrations is detailed below.

```{r tuna, echo=FALSE, out.width='100%'}
knitr::include_graphics("img/tuna.png")
```

@@

Saint Helena likely serves as a feeding ground for sub-adult fish, which migrate to spawning grounds upon reaching maturity (Wright et al., 2020).
Bigeye tuna (Thunnus obesus) and yellowfin tuna (Thunnus albacares) primarily spawn in tropical waters, such as the Gulf of Guinea and the Gulf of Mexico, during the summer (Carleton et al., 2010).
Recent tagging studies on yellowfin tuna show that they remain in Saint Helena waters for extended periods, ranging from six months to a year, during which they feed and grow (Wright et al., 2020).
They are typically recruited into the fishery between February and April each year and leave the population upon reaching a length of 120 cm (Wright et al., 2020).
Yellowfin tuna reach an average size of 96 to 120 cm at Cardno Seamount in May/June, suggesting migration out of the EEZ at this size (Wright et al., 2020; Wright et al., 2021).
Additionally, yellowfin tuna exhibit high fidelity to the area and remain in St. Helena waters for extended periods due to elevated productivity levels associated with the island and seamounts (White et al., 2007; Wright et al., 2021; Wright et al., 2019).

We utilized outputs from the SEAPODYM model, employed to characterize the movement dynamics of two tuna species in the Atlantic Ocean, as part of research conducted within the Mission Atlantic project [Merillet et al.
X (SOURCE)].
These outputs are based on the same four climate projections derived from NEMO-ERSEM physical outputs used in the StrathE2E driving data (see section X).
The projections are driven by two SSPs (SSP3-7.0 and SSP1-2.6) and two CMIP6 models (CNRM-CM6-1-HR and GFDL-ESM4).
Skipjack tuna (Katsuwonus pelamis) and albacore tuna (Thunnus alalunga) were modeled using SEAPODYM-MASS, an age-structured dynamic population model that accounts for prey availability.
Historical data were provided from 1990 to 2014, along with projections from 2015 to 2070, on a monthly basis and averaged across the Saint Helena model domain.
Data included biomass estimates for each SSP, model, and total stock biomass in the Atlantic.

To establish the baseline period from 2010-2019, historical data from 2010–2014 were combined with projections from 2015–2019.
The remaining time series were divided into decadal intervals from 2020–2029 to 2060–2069 for each of the four climate projections.

To identify the start and end days for annual immigration and emigration periods, we interpolated data to obtain daily values averaged by decade for each projection.
To synchronize with the StrathE2E annual cycle, each month was standardized to 30 days.
Seasonal cycles for each decade and projections were extracted using the decompose function in R, excluding trends and random components - for more information check the [function documentation](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/decompose/).
Immigration was defined as the period during which biomass increased from the 20th to the 80th percentile, while emigration was the period during which biomass decreased from the 80th to the 20th percentile.

```{r migration, echo=FALSE, fig.cap = "Seasonal component of migratory fish biomass derived from SEAPODYM model outputs. Points A to D indicate the start and end dates of immigration and emigration for each decade (2010–2019 to 2060–2069), under two climate scenarios (SSP1-2.6 and SSP3-7.0) and two forcing models (CNRM and GFDL).", out.width='400%'}
knitr::include_graphics("img/migration_Saint_Helena.png")
```

As the migratory guild in Saint Helena Island includes species beyond skipjack and albacore tunas, we utilized SEAPODYM outputs to parameterize the local biomass.
This was done using a scaling factor to account for the multi-species migratory biomass from the Atlantic, since local stock data are unavailable.
In addition to skipjack and albacore, we considered 16 other species contributing to migratory stocks in the Atlantic that may enter Saint Helena EEZ, including: wahoo (*Acanthocybium solandri*), frigate tuna (*Auxis thazard*), Galapagos shark (*Carcharhinus galapagensis*), common dolphinfish (*Coryphaena hippurus*), Atlantic sailfish (*Istiophorus albicans*), shortfin mako (*Isurus oxyrinchus*), white marlin (*Kajikia albida*), blue marlin (*Makaira nigricans*), blue shark (*Prionace glauca*), Atlantic bonito (*Sarda sarda*), smooth hammerhead (*Sphyrna zygaena*), yellowfin tuna (*Thunnus albacares*), southern bluefin tuna (*Thunnus maccoyii*), bigeye tuna (*Thunnus obesus*), Atlantic bluefin tuna (*Thunnus thynnus*), and swordfish (*Xiphias gladius*).

These migratory fish stocks in the Atlantic are managed under the International Convention for the Conservation of Atlantic Tunas (ICCAT), which relies on a consensus-driven decision-making process among member nations.
Biomass estimates for this study were derived from Sea Around Us (SAU) stock assessments, which utilize ICCAT data and the CMSY method (Pauly et al., 2020; see www.seaaroundus.org/cmsy-method for details).
These estimates covered the period from 2010 to 2018, and we used the averaged values for our calculations.
The combined biomass of the 18 key stocks in the Atlantic, which may enter the Ascension EEZ, totaled 5,503,844.45 tons.
This was treated as the ‘global’ stock biomass of migratory fish, with a portion assumed to enter the model domain in all decades and projections.

The scaling factor was derived by dividing this global biomass by the skipjack and albacore stocks provided by SEAPODYM.
This factor was then applied to local skipjack and albacore biomass from SEAPODYM to estimate the total migratory biomass for Saint Helena Island, averaged by decade, model, and SSP.
To quantify the proportion of Atlantic migratory biomass entering the model domain annually, we divided the scaled local biomass by the Atlantic migratory biomass.
To estimate the proportion of peak biomass that remained within the model domain without emigrating, we extracted the maximum and minimum values from the seasonal cycle.

Using these results, we derived the necessary parameters for lines 17–26 in the event_timing_SAINT_HELENA\_\*.csv files, for each decade and the four climate projections.

```{r tab_migrant_cap}

tab_migrantb_cap <- table_nums(name = "tab_migrantb", 
    caption = "Biological event timing parameters related to migratory fish constant across the time periods and climate projections")
```

`r table_nums("tab_migrantb")`

```{r tab_migrant_b, fig.cap = tab_migrantb_cap}

data_frame(Parameter = model[["data"]][["biological.events"]][["Description"]] %>% 
                        str_replace_all("_", " "),
           Value = model[["data"]][["biological.events"]][["Value"]]) %>% 
  .[17:20,] %>% 
  gt() %>%
  fmt_number(
    decimals = 3,
    use_seps = FALSE
  )
```

```{r migratory_baseline_cap}

migratory_baseline_cap <- table_nums(name = "migratory_baseline_cap", 
    caption = "Biological event timing parameters related to migratory fish for the baseline period (2010-2019) across the four climate projections.")
```

`r table_nums("migratory_baseline_cap")`

```{r migratory_baseline, fig.cap = migratory_baseline_cap}
migratory_fish_data <- data.frame(
  Parameter = c(
    "Propn of ocean popn entering domain each year",
    "Immigration start day",
    "Immigration end day",
    "Propn of peak popn in model domain which remains and does not emigrate",
    "Emigration start day",
    "Emigration end day"
  ),
  `SSP1-CNRM` = c("0.008001", "4", "145", "0.9590", "216", "291"),
  `SSP3-CNRM` = c("0.007891", "4", "110", "0.9608", "214", "291"),
  `SSP1-GFDL` = c("0.007757", "351", "107", "0.9466", "178", "280"),
  `SSP3-GFDL` = c("0.007892", "351", "109", "0.9553", "180", "280"),
  check.names = FALSE
)

migratory_fish_data %>%
  gt() %>%
  cols_align(align = "left", columns = vars(Parameter)) %>%
  cols_align(align = "center", columns = vars(`SSP1-CNRM`, `SSP3-CNRM`, `SSP1-GFDL`, `SSP3-GFDL`)) %>%
  cols_width(Parameter ~ px(220)) %>%
  tab_options(table.font.size = 15, table.align = "center")
```

@@

-   Temperature, vertical mixing coefficients, volume fluxes, and boundary nutrient, detritus and phytoplankton concentrations from outputs of an ensemble of NEMO-ERSEM coupled hydro-geochemical model runs with a 2015 historical/future split CITE YURI XXX.
-   Surface shortwave radiation from CMIP6 used to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
-   Freshwater volume outflows from CMIP6 used to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
-   River nitrate and ammonia concentrations taken from the IMAGE model to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
-   Atmospheric deposition of nitrate and ammonia from ISIMIP3a ([Inter-Sectoral Impact Model Intercomparison Project](https://data.isimip.org/10.48364/ISIMIP.759077.2); [@yang22]) with SSP projects from ISIMIP3b used to force the NEMO-ERSEM coupled hydro-geochemical runs mentioned above.
-   Remote sensing data products on Suspended Particulate Matter (Globcolour L3b; <ftp://ftp.hermes.acri.fr/GLOB/merged/month/>).
-   Wave height, period, and direction from the ERA-5 reanalysis monthly means accessed through [CDS](https://cds.climate.copernicus.eu/cdsapp#!/dataset/10.24381/cds.f17050d7?tab=overview) for 1980-2021 [@Hersbach19].

Details of how these data were processed are given below, supported by the [nemomedusR](https://jack-h-laverick.github.io/nemoRsem/) and [MiMeMo.tools](https://jack-h-laverick.github.io/MiMeMo.tools/) packages.

### NEMO-ERSEM ensemble: {.unnumbered}

Four different NEMO-ERSEM runs were used to parameterise different versions of the Ascension Island implementation of StrathE2E.
These four runs are a 2x2 factorial cross of two future projected scenarios (SSP126 and SSP370) with a historical hindcast from 2015, and forcing by two atmospheric models (CNRM and GFDL).
In the following sections model output was processed for a 2010-2019 baseline period, and then decadal projections from 2020-2029 until 2060-2069.

### Vertical mixing coefficients between the model layers: {.unnumbered}

Vertical diffusivity from the NEMO-ERSEM coupled hydro-geochemical model output was interpolated for each grid cell at the `r SDepth`m boundary depth between the shallow and deep layers of the offshore zone, and the `r DDepth`m boundary at the deep sea overhang.
These values were summarised as monthly averages into period-specific climatological annual cycles of data for decadal periods for all combinations of SSPs and forcings.

@@

Derived by monthly averaging values at grid points within the inshore and vertical layers of the offshore zones from the NEMO-ERSEM coupled hydro-geochemical model output, weighted by grid point volumes.
Values were summarised into period-specific climatological annual cycles of data.

@@

Derived from HadGEM2-ES model output [@jones11] which forces the NEMO-ERSEM model run used throughout our implementation.
Monthly mean values were summarised into climatological annual cycles of data for decadal periods and NEMO-ERSEM ensembles.

@@

Sourced from the "histsoc" files for a 1901 - 2021 hindcast as monthly averages [@yang22], available from [CDS](https://cds.climate.copernicus.eu/cdsapp#!/dataset/10.24381/cds.f17050d7?tab=overview).
Monthly values were summarised into climatological annual cycles of monthly oxidised and reduced nitrogen deposition rates extracted for 2010-2019.
ISIMIPb Projections for SSP126 and SSP370 were processed in the same way for decadal periods up to and including the 2060s.

@@

[old information?]
Freshwater inflow derived from HadGEM2-ES model output [@jones11] which forces the NEMO-MEDUSA model run used throughout our implementation.
Monthly values were summaries into a climatological annual cycle of data for both the 2010-2019 and 2040-2049 periods.

@@

Monthly averaged daily inflow and outflow volume fluxes derived by integrating daily mean velocities directed perpendicular to transects along the model domain boundary at grid points in each depth layer along transects through outputs from the NEMO-ERSEM coupled hydro-geochemical model output.
Monthly averaged daily inflow volume fluxes then divided by the volume of the receiving layer in the model domain to estimate a daily flushing rate.

@@

NEMO-ERSEM outputs included nitrate, ammonia, phytoplankton and suspended detritus.
We calculated the depth-averaged concentrations for pixels within the shallow and deep layers of StrathE2E.
We then sampled the pixels using the same transects around the model domain as for sampling volume fluxes.
Only transects where water flowed into the model domain were sampled, and the average concentration of inflowing waters for target variables was calculated weighting by the flow rate across a transect and the cross-sectional area represented by a transect (average depth and length).
Concentrations were then averaged into climatological annual cycles.
The flow weighted averaging of concentrations was calculated on the 5-day NEMO-ERSEM timestep, before averaging to a monthly climatology of concentrations.

@@

```{r pole-and-line, echo=FALSE, out.width = '100%'}
knitr::include_graphics("img/pole-and-line.png")
```

@@

```{r rock-fishing, echo=FALSE, out.width = '100%'}
knitr::include_graphics("img/rock-fishing.png")
```

@@

```{r fleets, echo=FALSE, fig.cap="Fishing fleets defined for the StrathE2E Saint Helena EEZ model, their operational descriptions, target-species and discards levels.", out.width = '120%'}
knitr::include_graphics("img/Fleets.png")
```

@@

effort_tab <- effort_table %>%
  gt() %>%
  tab_options(
    table.font.size = 15,
    column_labels.font.weight = "bolder"
  ) %>%
  cols_align(align = "center") %>%
  cols_label(
    Fleets = "Fleets StrathE2E Saint Helena",
    `Effort (s.year-1)` = html("Effort (s.year<sup>-1</sup>)"),
    `Effort (s.m2-1.day-1)` = html("Effort (s.m<sup>2</sup>.day<sup>-1</sup>)"),
  )

@@

The distribution of each fleet across habitats was assessed by integrating literature data with sediment habitat information. The majority of fleets operate offshore within the overhang zone of the model. Exceptions include the artisanal handlines and hand-gathering and pot fishing fleets, which are predominantly concentrated in the inshore zone. Handline fishing for grouper and pot fishing for lobsters are confined to the inshore rock and gravel habitats area and are not present around the Bonaparte and Cardno seamounts (Collins et al., 2012; Rees et al., 2016; Riley et al., 2020). Baitfish (Decapterus spp.) are caught with nets at the surface in waters 10 to 20 meters deep (Carleton et al., 2010), usually near sandy and rocky inshore beaches (Edwards, 1990). We allocated the fishing effort per habitat as outlined in Table X.

```{r tab_distribution_cap}

tab_distribution_cap <- table_nums(name = "tab_distribution", 
    caption = "Spatial distribution of fishing activity across seabed habitat classes, inferred for Saint Helena based on the characteristics of local fishing fleets. The 'deep ocean' habitat refers to the portion of the offshore zone that extends into deep ocean waters, without a seafloor.")
```

`r table_nums("tab_distribution")`

```{r tab_distribution, fig.cap = tab_distribution_cap}

distribution_tab <- data.frame(
  Fleets = c("Pole and line", "Longlines", "Handlines", "Handlines and dip-nets (bait)", "Pots and hand gathering"),
  `Shallow_rock` = c(0, 0, 0.5, 0.25, 0.5),
  `Shallow_sand` = c(0, 0, 0, 0.5, 0),
  `Shallow_gravel` = c(0, 0, 0.5, 0.25, 0.5),
  `Deep_ocean` = c(1, 1, 0, 0, 0)
)

distribution_tab <- distribution_tab %>%
  gt() %>%
  tab_options(table.font.size = 15, column_labels.font.weight = "bolder") %>%
  cols_align(align = "center") %>%
  cols_label(`Shallow_rock` = "Shallow rock",
             `Shallow_sand` = "Shallow sand",
             `Shallow_gravel` = "Shallow gravel",
             `Deep_ocean` = "Deep ocean",
             Fleets = "Fleets StrathE2E Saint Helena")

distribution_tab
```

@@

Calculating the production involved compiling and adjusting biomass values for the model area and converting them to molar nitrogen mass following @greenstreet96.
Subsequently, we convert the biomass for each species into gross production using PB ratios and converted them to molar nitrogen mass following @greenstreet96.
The PB ratios sourced was also used as target data.
For top-predators, net production was calculated using the relationship net production = 0.6 \* gross production.

@@

tab <- production_tab %>%
  gt() %>%
  fmt_number(
    columns = c(`P/B`, `mMN convertion`, `Production(t)`),
    decimals = 2,
    use_seps = FALSE
  ) %>%
  fmt_number(
    columns = c(`Production (mMN.m-2)`),
    decimals = 6,
    use_seps = FALSE
  ) %>%
  tab_options(
    table.font.size = 15,
    column_labels.font.weight = "bolder",
  ) %>%
  cols_align(
    align = "center"
  ) %>%
  cols_label(
    Guild = "Guild",
    `P/B` = "P/B",
    `Production(t)` = "Production(t)",
    `mMN convertion` = "mMN convertion",
    `Production (mMN.m-2)` = html("Production (mMN.m<sup>-2</sup>)")
  )
tab
```

@@

By multiplying the estimated number of individuals by the average weight of each species, as provided by BirdLife International (2024) and Sealife (2024), we obtained an approximate seabird biomass of 9.82 tons.
To obtain the production per year we used the PB ratio for seabirds from the Ecopath model of Capitani et al. (2021) based on Rocas Atoll, another tropical reef ecosystem in the Southwest Atlantic.

```{r seabirds, echo=FALSE, out.width = '100%'}
knitr::include_graphics("img/seabirds.png")
```

@@

```{r whale, echo=FALSE, out.width = '100%'}
knitr::include_graphics("img/whale.png")
```

@@

```{r lobster, echo=FALSE, out.width = '100%'}
knitr::include_graphics("img/lobster.png")
```

@@

## Diet proportions {.unnumbered}

@@

@@

tab <- diet_tab %>%
  gt() %>%
  tab_options(
    table.font.size = 15,
    column_labels.font.weight = "bolder",
  ) %>%
  cols_align(
    align = "center"
  ) 
tab

@@
